Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}

### 2017-09-29 paper/casestudyEPlus.tex
Modified introduction. First complete draft.
Added flow scheme, motivation. Implemented comments from Italy.

Fig.~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Sec.~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

\begin{figure}[h!]
	\centering
	\hspace{-0.3cm}\includegraphics[width=22pc]{figures/eplus_validation.eps}	
	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances. The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
	\label{F:sepvars}
	\captionsetup{justification=centering}
\end{figure}

@@

The results are shown in Fig.~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Fig.~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%. The difference between the predicted power consumption and that in the closed-loop simulation in Fig.~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}. Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

>>>>

Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

\begin{figure}[h!]
	\centering
	\includegraphics[width=24pc]{figures/eplus_validation.eps}	
	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances. The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
	\label{F:sepvars}
	\captionsetup{justification=centering}
\end{figure}

@@

The results are shown in Figure~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%. The difference between the predicted power consumption and that in the closed-loop simulation in Figure~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}. Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2017-09-29 paper/casestudyReal.tex
awesome figure
Added flow scheme, motivation. Implemented comments from Italy.

\section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
\label{S:realCaseStudy}

In this section we show an application of DPC to a real case study. In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus models of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.

\subsection{Description of the house}\label{SS:descriptionHouse}
The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}. The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor, that includes $4$ rooms, and an attic without heating system. Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.

@@

Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building. Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and as well for water supply, has a complete independence from the utilities.   
The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance. The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=27pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of the four differently oriented rooms of the house as shown in Figure \ref{F:houseFloors}. The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy. The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
\begin{figure}[h!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}

The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption of the house. These models will be used for performance comparison of DPC with respect to a classical bang-bang controller. In particular DPC will be set up to provide an optimal ON/OFF scheduling policy for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. To this aim, we also create random forest models for power consumption and room temperatures to be used in the closed-loop simulations.

@@

To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed. The model has been created according to the K\"{o}ppen-Geiger climate classification that, based weather data of L'Aquila shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}, classifies Italy in Csa, Cfa and Csb climate zones, with a warm temperate climate [4], .

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\subfigure[Thermal energy.]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models} For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tP^{r}_{\mathrm{k+1|k}},\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\\\tT^{3}_{\mathrm{k+j|k}},\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests. In each set we have a model for the power consumption of the house ($\tP^{r}$ and $\tP$) and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms ($\tT^{r_i}$ and $\tT^i$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

@@

\subsection{DPC and bang-bang controllers} We set up $2$ different controllers, DPC and bang-bang controller, to obtain a scheduling policy to switch the radiators ON and OFF in order to keep the temperature of room $3$ within a comfort range. Since the heating system serves all of the $4$ rooms simultaneously, without giving the possibility to control the temperature of each room independently, we setup the problem defining the comfort range only for one room. Other rooms temperatures will follow the scheduling policy. Room $3$ has been chosen randomly. In the following we describe the $2$ controllers.

@@

%\begin{figure}[h!]
%	\begin{center}
%		\includegraphics[width=26pc]{figures/Energy_all.eps}
%	\end{center}
%	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
%	\label{F:comparison_all_energy}
%\end{figure}
%An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}. Results are shown in Figure \ref{F:comparison_all_energy_E+}.
In Figure \ref{F:comparison_all_energy_E+}, using the thermal energy consumption model derived in Section \ref{SS:energyPlusmodel} using EnergyPlus, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy. In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the testing data, i.e. $15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that correspond to the $25.4\%$ over a period of $15$ days. In case of medium violations get an energy saving of $57\, kWh$, that is $32.2\%$. Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that correspond to the $49.2\%$.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
	\end{center}
	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
	\label{F:comparison_all_energy_E+}
\end{figure}

>>>>

\section{CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
\label{S:realCaseStudy}

In this section we show an application of DPC to a real case study. The section organization is shown in Figure \ref{F:overview}. More precisely, we consider a real house located in L'Aquila, Italy, as a case study that we describe in Section \ref{SS:descriptionHouse}. Using historical data from this house, random forest models are built in Section \ref{SS:randomForestsModels} to be used either as plant or as predictive models for DPC. In Section \ref{SS:controllersDPCandBangbang}, DPC and bang-bang control strategies are defined to be used in the close-loop simulations. Finally in Section \ref{SS:simulationResults} controllers performance are compared in terms of energy saving using a thermal energy model of the house built via EnergyPlus in Section \ref{SS:energyPlusmodel}, using both historical data, and house construction layout and materials.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Section \ref{S:realCaseStudy} organization scheme.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\subsection{Description of the house}\label{SS:descriptionHouse}
The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}. The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system. Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.

@@

Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building. Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, has a complete independence from the utilities.   
The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance. The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=27pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$). The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy. The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
\begin{figure}[h!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}

The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption assessment of the house. These models will be used for performance comparison of DPC with respect to a classical bang-bang controller. In particular DPC will be set up to provide an optimal ON/OFF scheduling policy for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. To this aim, we also create random forest models for power consumption and room temperatures to be used in the closed-loop simulations.

@@

To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed. The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}. According to the K\"{o}ppen-Geiger climate classification, Italy is classified in Csa, Cfa and Csb climate zones, with a warm temperate climate \cite{Peel2007}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila, Italy.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\subfigure[Thermal energy.]{
			\label{F:houseComparisonExperimentalEnergy}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
		}
		\subfigure[Thermal power.]{
			\label{F:houseComparisonExperimentalPower}
			\centering
			\includegraphics[width=24pc]{figures/Thermal_power.eps}
		}
	\end{center}
	\caption{Comparison between numerical and experimental data.}
	\captionsetup{justification=centering}
	\label{F:houseComparisonExperimental}
\end{figure}
\subsection{Random forest models}
\label{SS:randomForestsModels}
For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests. In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors. In $S_2$ we have a model for the power consumption of the house ($\tP$). Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house. For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$. Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}. The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms). The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$). All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$. All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).

@@

\subsection{DPC and bang-bang controllers}
\label{SS:controllersDPCandBangbang}
We set up $2$ different controllers, DPC and bang-bang controller, to obtain a scheduling policy to switch the radiators ON and OFF in order to keep the temperature of room $3$, i.e. the living room, within a comfort range. Since the heating system serves all of the $4$ rooms simultaneously, without giving the possibility to control the temperature of each room independently, we setup the problem defining the comfort range only for one room. Other rooms temperatures will follow the scheduling policy. Room $3$ has been chosen randomly. In the following we describe the $2$ controllers.

@@

%\begin{figure}[h!]
%	\begin{center}
%		\includegraphics[width=26pc]{figures/Energy_all.eps}
%	\end{center}
%	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
%	\label{F:comparison_all_energy}
%\end{figure}
%An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}. Results are shown in Figure \ref{F:comparison_all_energy_E+}.
In Figure \ref{F:comparison_all_energy_E+}, using the thermal energy consumption model derived in Section \ref{SS:energyPlusmodel} using EnergyPlus, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy. In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the testing data, i.e. $15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. We can see that to the bang-bang control strategy is associated an energy consumption approximately equal to $177\,kWh$. If we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that corresponds to the $25.4\%$ over a period of $15$ days. In case of medium violations get an energy saving of $57\, kWh$, that is $32.2\%$. Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that corresponds to the $49.2\%$.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
	\end{center}
	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
	\label{F:comparison_all_energy_E+}
\end{figure}

### 2017-09-29 paper/cs-comparison.tex
completed section 5 and fixed figures problems
Modified introduction. First complete draft.
Added flow scheme, motivation. Implemented comments from Italy.

The performance is compared for 3 days in winter, i.e. January 28-31 and 3 days in summer, i.e. May 1-3. These are shown on the same plots in Fig.~\ref{F:comparison}.
The sampling time in the simulations is 1 hr. The control horizon $N$ and the order of autoregression are both 6 hrs. The training procedure required a few minutes in the case of trees and 2 hrs for forests on a Win 10 machine with an i7 processor and 8GB memory.
The cooling usage factor $\mathsf{C}$ is constrained in $[0,1]$, the heat input in $[0,23]\ \mathrm{W/m^2}$, and the room temperature in $[19,25]\ \mathrm{^oC}$ during the winters and $[20,26]\ \mathrm{^oC}$ during the summers.
The optimization is solved using CPLEX \cite{IBM}.

The external disturbances - solar gains, internal gain due to equipment and dry-bulb temperature during the chosen periods are shown in Fig.~\ref{F:dist}. The internal gain due to occupancy was proportional to the gain due to equipment. 
The reference temperature is chosen to be 22 $\mathrm{^oC}$. Due to cold weather, which is evident from the dry-bulb temperature, the heater is switched on during the night to maintain the thermal comfort requirements. When the building is occupied during the day, due to excessive internal gains, the building requires cooling. The lighting in the building is adjusted to meet the minimum light requirements.
The optimal cooling usage factor and the radiator power for MPC, DPC-En and DPC-RT are shown in Fig.~\ref{F:control1} and Fig.~\ref{F:control2}, respectively. The control strategy with DPC-En shows a remarkable similarity to MPC, switching on/off the equipments at the same time with similar usage. However, the performance with DPC-RT is much different and worse. DPC-RT inherently suffers from high variance which is also evident in the control strategy, thus making it unsuitable for practical purposes. 
Although it seems like that adding the rate constraints to DPC-En would smoothen its behavior, this was avoided because the sampling time of the system is 1 hr which is already too high. The room temperature profile in Fig.~\ref{F:state} is close to the reference in the case of DPC-En as well as MPC. 
Fig.~\ref{F:obj} shows that the cumulative cost of the objective function is, as expected, minimum for MPC, and a bit higher for DPC-En. The cost for DPC-RT blows up around 12 noon on 30$^{th}$ January as one of the slack variables is non-zero, which happens due to high model inaccuracy.

The quantitative performance comparison is shown in Tab.~\ref{T:comparison}. MPC tracks the reference more closely at the expense of higher input costs in comparison to DPC-En. The higher cost of the inputs in MPC is also due to lighting. DPC-En explains 70.1\% variation in the optimal control strategies obtained from MPC while DPC-RT explains only 1.8\%. The mean optimal cost of DPC-En is more than MPC, and is maximum for DPC-RT due to a constraint violation.

\begin{table}[h!]
	\centering
	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
	\captionsetup{justification=centering}
	\begin{tabular}{c|c|c|c|c}
		\toprule
		& explained & mean objective& mean input  & mean  \\
		&  variance$[\mathrm{-}]$ & value $[\mathrm{-}]$ & cost $[-]$ & deviance $[\mathrm{^oC}]$ \\     
		\midrule
		MPC    &  $\mathrm{-}$ &  22.60 & 17.16  &  0.26  \\
		DPC-En   & 70.1\% &  39.26  & 15.12 &  0.48 \\
		DPC-RT  & 1.8\% & 204.55 & 16.84 &  0.57 \\
		\bottomrule
	\end{tabular}
	\label{T:comparison}
\end{table}
\begin{figure}[h!]
	\subfigure[External disturances: solar gains, internal gain due to equipment and dry-bulb temperature.]{
		\label{F:dist}
		\centering
		\includegraphics[width=20pc]{figures/disturbances.eps}
	}
	\subfigure[Optimal control input: cooling usage factor $\mathsf{C}$ with  $0 \leq \mathsf{C} \leq 1$. DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control1}
		\centering
		\includegraphics[width=20pc]{figures/input3.eps}
	}
	\subfigure[Optimal control input: radiator heat $\mathsf{H}$ with  $0 \leq \mathsf{H} \leq 23 \ \mathrm{W/m^2}$. Again, DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control2}
		\includegraphics[width=20pc]{figures/input4.eps}
	}
	\subfigure[Room temperature has time varying bounds. When the building is occuped the constraints are relaxed, else $19(20) \leq \mathsf{T}_{\mathrm{in}} \leq 25(26) \ \mathrm{^oC}$ in January(May). MPC and DPC-En are able to track the reference temperature ($22 \ \mathrm{^oC}$) closely.]{
		\label{F:state}
		\centering
		\includegraphics[width=20pc]{figures/state.eps}
	}
	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
		\label{F:obj}
		\centering
		\includegraphics[width=20pc]{figures/cumsumcost.eps}
	}
	%\vspace{-1cm}
	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
	\label{F:comparison}
	\captionsetup{justification=centering}
\end{figure}

Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.
However, one major limitation of the bilinear model is that the information about the building power consumption is not available. Much nonlinearities in the system are due to equipment efficiencies which are not considered in the bilinear case but are very important for practical purposes.

>>>>

The performance is compared for 3 days in winter, i.e. January 28-31 and 3 days in summer, i.e. May 1-3. These are shown on the same plots in Figure~\ref{F:comparison}.
The sampling time in the simulations is 1 hr. The control horizon $N$ and the order of autoregression are both 6 hrs. The training procedure required a few minutes in the case of trees and 2 hrs for forests on a Win 10 machine with an i7 processor and 8GB memory.
The cooling usage factor $\mathsf{C}$ is constrained in $[0,1]$, the heat input in $[0,23]\ \mathrm{W/m^2}$, and the room temperature in $[19,25]\ \mathrm{^oC}$ during the winters and $[20,26]\ \mathrm{^oC}$ during the summers.
The optimization is solved using CPLEX \cite{IBM}.

The external disturbances - solar gain, internal gain due to equipment and dry-bulb temperature during the chosen periods are shown in Figure~\ref{F:dist}. The internal gain due to occupancy was proportional to the gain due to equipment. 
The reference temperature is chosen to be 22 $\mathrm{^oC}$. Due to cold weather, which is evident from the dry-bulb temperature, the heating system is switched on during the night to maintain the thermal comfort requirements. When the building is occupied during the day, due to excessive internal gain, the building requires cooling. The lighting in the building is adjusted to meet the minimum light requirements.
The optimal cooling usage factor and the radiator power for MPC, DPC-En and DPC-RT are shown in Figure~\ref{F:control1} and Figure~\ref{F:control2}, respectively. The control strategy with DPC-En shows a remarkable similarity to MPC, switching on/off the equipments at the same time with similar usage. However, the performance with DPC-RT is much different and worse. DPC-RT inherently suffers from high variance which is also evident in the control strategy, thus making it unsuitable for practical purposes. 
Although it seems like that adding the rate constraints to DPC-En would smoothen its behavior, this was avoided because the sampling time of the system is 1 hr which is already too high. The room temperature profile in Figure~\ref{F:state} is close to the reference in the case of DPC-En as well as MPC. 
Figure~\ref{F:obj} shows that the cumulative cost of the objective function is, as expected, minimum for MPC, and a bit higher for DPC-En. The cost for DPC-RT blows up around 12 noon on 30$^{th}$ January as one of the slack variables is non-zero, which happens due to high model inaccuracy.

The quantitative performance comparison is shown in Table~\ref{T:comparison}. MPC tracks the reference more closely at the expense of higher input costs in comparison to DPC-En. The higher cost of the inputs in MPC is also due to lighting. DPC-En explains 70.1\% variation in the optimal control strategies obtained from MPC while DPC-RT explains only 1.8\%. The mean optimal cost of DPC-En is more than MPC, and is maximum for DPC-RT due to a constraint violation.

\begin{table}[h!]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		& explained & mean objective& mean input  & mean  \\
		&  variance$[\mathrm{-}]$ & value $[\mathrm{-}]$ & cost $[-]$ & deviance $[\mathrm{^oC}]$ \\     
		\midrule
		MPC    &  $\mathrm{-}$ &  22.60 & 17.16  &  0.26  \\
		DPC-En   & 70.1\% &  39.26  & 15.12 &  0.48 \\
		DPC-RT  & 1.8\% & 204.55 & 16.84 &  0.57 \\
		\bottomrule
	\end{tabular}
	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
	\captionsetup{justification=centering}
	\label{T:comparison}
\end{table}

\begin{figure}[t!]
	\begin{center}
	\subfigure[External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.]{
		\label{F:dist}
		\centering
		\includegraphics[width=16pc]{figures/disturbances.eps}
	}
	\subfigure[Optimal control input: cooling usage factor $\mathsf{C}$ with  $0 \leq \mathsf{C} \leq 1$. DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control1}
		\centering
		\includegraphics[width=16pc]{figures/input3.eps}
	}
	\subfigure[Optimal control input: radiator heat $\mathsf{H}$ with  $0 \leq \mathsf{H} \leq 23 \ \mathrm{W/m^2}$. Again, DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control2}
		\includegraphics[width=16pc]{figures/input4.eps}
	}
	\subfigure[Room temperature has time varying bounds. When the building is occuped the constraints are relaxed, else $19(20) \leq \mathsf{T}_{\mathrm{in}} \leq 25(26) \ \mathrm{^oC}$ in January(May). MPC and DPC-En are able to track the reference temperature ($22 \ \mathrm{^oC}$) closely.]{
		\label{F:state}
		\centering
		\includegraphics[width=16pc]{figures/state.eps}
	}
\end{center}
\end{figure}
\begin{figure}[h!]
\begin{center}
	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
		\label{F:obj}
		\centering
		\includegraphics[width=16pc]{figures/cumsumcost.eps}
	}
	\end{center}
	%\vspace{-1cm}
	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
	\label{F:comparison}
	\captionsetup{justification=centering}
\end{figure}
Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.
However, one major limitation of the bilinear model is that the information about the building power consumption is not available. Much nonlinearities in the system are due to equipment efficiencies which are not considered in the bilinear case but are very important for practical purposes.
